---
title: "CSN - Second Lab"
author: "Kymry Burwell, Laura Cebollero"
geometry: "left=3cm,right=3cm,top=2cm,bottom=2cm"
date: "October 7th, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
requiredPackages <- c("igraph", "ggplot2", "data.table", "expss", "stats4", "VGAM", "knitr", "rmarkdown")
for (pac in requiredPackages) {
  if(!require(pac,  character.only=TRUE)){
    install.packages(pac, repos="http://cran.rstudio.com")
    library(pac,  character.only=TRUE)
  } 
}
rm(pac)
rm(requiredPackages)

# Set WD and load data
wd = getwd()
if(grepl("nora", wd)) {
    setwd("~/git/csn-labs/02")
} else {
  setwd("~/Google Drive/UPC/Fall 2018/CSN/Labs/git_labs/Complex-and-Social-Networks/02")
}
rm(wd)
```
```{r metric_functions, include=FALSE}

get_H <- function(k, gamma) {
  return(sum(seq(1, k)^(-gamma)))
}

get_AIC <- function(m2logL,K,N) {
  m2logL + 2*K*N/(N-K-1) # AIC with a correction for sample size
}

get_MP <- function(x) { return(sum(log(x)))}

get_C <- function(x) { C = 0
  for(i in x){
    j = 2
    while( j <= i){
      C <- C + log(j)
      j = j+1
    }
  } 
  return(C)
}
```

```{r comp_functions, include=FALSE}


############# Creating summary table

write_summary <- function(language,file) {
    degree_sequence = read.table(file, header = FALSE)
    l <- list(language,length(degree_sequence$V1),sum(degree_sequence$V1),max(degree_sequence$V1),sum(degree_sequence$V1)/length(degree_sequence$V1),length(degree_sequence$V1)/sum(degree_sequence$V1), get_MP(degree_sequence$V1), get_C(degree_sequence$V1))
    return(l)
}


create_sum_table <- function(source){
    
    temp_df <- data.table("Language" = character(),
                          "N" = numeric(),
                          "M" = numeric(),
                          "Maximum Degree" = numeric(),
                          "M/N" = numeric(),
                          "N/M" = numeric(),
                          "MP" = numeric(),
                          "C" = numeric(),
                          stringsAsFactors = FALSE)
    
    for (x in 1:nrow(source)){
        file <- source$file[x]
        language <- source$language[x]
        sequence <- write_summary(language, file)
        temp_df <- rbind(temp_df,sequence)
    }
    return(temp_df)
}               


###########  Estimating log-likelihood parameters ########### 
compute_log_likelihoods <- function(M, N, maxDegree, MP, C){

    # Displaced Poisson function
    minus_log_likelihood_poisson <- function(lambda){
      -(M * log(lambda) - N*(lambda + log(1 - exp(-lambda))) - C) 
    }
    
    # Displaced Geometric function
    minus_log_likelihood_geometric <- function(q){
      -(M - N) * log(1 - q) - N * log(q)
    }
    
    # Minus log-likelihood function (minus_log_likelihood = -L)
    minus_log_likelihood_zeta <- function(gamma){ 
      gamma * MP + N * log(zeta(gamma))
    }
    
    # Minus log-likelihood zeta function with lambda - 2
    minus_log_likelihood_zeta2 <- function(){
      2 * MP + N * log(pi^2/6)
    }
    
    # Minus log-likelihood right-truncated zeta
    minus_log_likelihood_zeta3 <- function(gamma){
      gamma * MP + N * log(get_H(maxDegree, gamma))
    }
    
    # Altmann function
    minus_log_likelihood_altmann <- function(lambda, delta){
      c * maxDegree^(-lambda) * e^(-delta * maxDegree)
    }
    
    mle_poisson <- mle(minus_log_likelihood_poisson,
                    start = list(lambda = M/N),
                    method = "L-BFGS-B",
                    lower = c(1.000001))

    mle_geometric <- mle(minus_log_likelihood_geometric,
                    start = list(q = N/M), 
                    method = "L-BFGS-B",
                    lower = c(0.01), 
                    upper = c(.99)) # No higher than .99 (ex: .999)

    mle_zeta <- mle(minus_log_likelihood_zeta,
                    start = list(gamma = 2),
                    method = "L-BFGS-B",
                    lower = c(1.00000001))

    mle_zeta2 <- mle(minus_log_likelihood_zeta,
                    fixed = list(gamma = 2),
                    method = "L-BFGS-B",
                    lower = c(1.00000001))

    # TODO: Check warning below. Doesn't look good.
    # Update 6/10: Check there is still warnings. Was using length(x) instead of N (wrong)
    mle_zeta3 <- mle(minus_log_likelihood_zeta3,
                    start = list(gamma = 2), #Alternatively, k = N
                    method = "L-BFGS-B",
                    lower = c(1.001, N))

    
    vec <- c(
        attributes(summary(mle_poisson))$coef[1],
        attributes(summary(mle_geometric))$coef[1],
        attributes(summary(mle_zeta))$coef[1],
        2,
        attributes(summary(mle_zeta3))$coef[1],
        get_AIC(attributes(summary(mle_poisson))$m2logL, 1, N),
        get_AIC(attributes(summary(mle_geometric))$m2logL, 1, N),
        get_AIC(attributes(summary(mle_zeta))$m2logL, 1, N),
        get_AIC(attributes(summary(mle_zeta2))$m2logL, 1, N),
        get_AIC(attributes(summary(mle_zeta3))$m2logL, 1, N)
    )

    return(vec)
}



########### MODEL SELECTION ########### 
compute_coeffs_table <- function(summary_table) {
    coeff_table <- data.table("Language" = character(),
                          "lambda" = numeric(),
                          "q" = numeric(),
                          "gamma 1" = numeric(),
                          "gamma 2" = numeric(),
                          "k_max" = numeric(),
                          stringsAsFactors = FALSE)
    
    aic_table <- data.table("Language" = character(),
                          "Poisson" = numeric(),
                          "Geometric" = numeric(),
                          "Zeta gamma = 2" = numeric(),
                          "Zeta" = numeric(),
                          "RT Zeta" = numeric(),
                          stringsAsFactors = FALSE)
    
    for (i in seq(length(summary_table$Language))) {
        language <- summary_table[i]$Language 
        M <- summary_table[i]$M
        N <- summary_table[i]$N
        MP <- summary_table[i]$MP
        C <- summary_table[i]$C
        maxDegree <- summary_table[i]$`Maximum Degree`
        resultList <- compute_log_likelihoods(M, N, maxDegree, MP, C)
        g <- as.list(resultList[1:5])
        h <- as.list(resultList[6:10])
        coeff_table <- rbind(coeff_table, c(language, g))
        aic_table <- rbind(aic_table, c(language, h))
    }
    
    # Computing delta: AIC - best AIC 
    for (i in seq(length(aic_table$'Poisson'))) {
        aic_table[i, 2:6] <- aic_table[i, 2:6] - min(aic_table[i, 2:6])
    }
    return(list(coeff_table, aic_table))
}

```
```{r english_sequence, echo=FALSE}
    degree_sequence = read.table("./data/English_out-degree_sequence.txt",
                                 header = FALSE)
    num_nodes <- nrow(degree_sequence)
    sum_degree <- sum(degree_sequence)
    mean_degree <- sum(degree_sequence)/dim(degree_sequence)[1]
    
    # Remove nodes of out-degree 0 (k=0)
    degree_sequence$V1 <- degree_sequence[!(degree_sequence$V1==0),]
    
    # Barplots of data
    degree_sequence = read.table("./data/English_out-degree_sequence.txt", header = FALSE)
    degree_spectrum = table(degree_sequence)
    x <- degree_sequence$V1
    M <- sum(x)
    N <- length(x)
```
# Introduction
The aim of this lab project is to analyze a degree distribution and select a theoretic model that best fits it. There are three sequences of which we can work on:

1. Undirected degree sequence.
2. In-degree sequence.
3. Out-degree sequence.

In our case, we have chosen to work with the out-degree one  for 10 different languages.

The distributions we will be testing are the following:

- Poisson distribution ($\lambda$ parameter)
- Geometric distribution ($q$ parameter)
- Zeta distribution ($\gamma$ parameter)
- Zeta distribution ($\gamma = 2$ parameter)
- Right truncated distribution ($\gamma$ and $k_{max}$ parameteres)
- Altmann distribution ($k_{max}$, $\delta$, and $\gamma$ parameters)

The first step on the analysis is to compute different metrics for each language, such as the length of the sequence (N) and the maximum degree, among others.

Additionally, to make our computations easier we have added a couple of metrics that we were not required in Table 1, which are MP and C.

```{r include=FALSE, echo=FALSE}

out_source = read.table("list_out.txt", header = TRUE, as.is = c("language","file")) 
(summary_table <- create_sum_table(out_source))

```

The resulting table is the following:

```{r echo=FALSE}
kable(summary_table)
```

In the table above, N represents the number of nodes in the network, M is the sum of degrees of all nodes, Maximum Degree is the highest degree in the give language, M/N is the average degree, N/M is the inverse of the average degree, MP is the sum of the log of degrees, and C is the the following $\sum_{i=1}^{N} \sum_{j=2}^{k_i} log(j)$.

Next, we will look at a few bar plots for the English langauge to get a visual idea of the degree distribution. 

```{r barplot, echo=FALSE, out.width="250px", out.height="250px", fig.align="center"}
 barplot(degree_spectrum, main = "English", xlab = "out-degree", ylab = "number of vertices")
 barplot(degree_spectrum, main = "English", xlab = "out-degree", ylab = "number of vertices", log = "xy")
```

We can see from the above plots that nodes with small out-degree are more common than nodes with high out-degree.

# Results
Having computed the basic metrics, we now proceed to compute the most likely parameters for the various distributions. To do this, we are trying to find the parameters that minimize the minus log-likelihood function. To help expedite the process, we begin with default parameters, which act as our best initial guess. Theses consist of the following:

- $\lambda_{0}$ = M/N
- $q_{0}$ = N/M
- $\gamma_{0}$ = 2
- $k_{max, 0}$ = N


```{r comp_tables, include=FALSE, echo=FALSE}
aic_c_table <- compute_coeffs_table(summary_table)

coeffs_table <- as.data.table(aic_c_table[1])
aic_table <- as.data.table(aic_c_table[2])

```

```{r echo=FALSE}
kable(coeffs_table)
```

In order to verify the methods we used to compute the minus log-likelihood parameters, we will test them on contrived data where the distribution is known a priori. We have 8 different data sets that were created using geometric and zeta distributions of varying parameters. 

## Geometric test

For the different instances of the geometric test table, we have computed first the coefficients table:

```{r echo=FALSE}

geometric_01 = read.table("./samples_from_discrete_distributions/data/sample_of_geometric_with_parameter_0.1.txt",
                             header = FALSE)


geom_source = read.table("./list_geometric.txt", header = TRUE, as.is = c("language","file")) # For testing 

test_table <- create_sum_table(geom_source)
aic_c_table_test <- compute_coeffs_table(test_table)

coeffs_table_test <- as.data.table(aic_c_table_test[1])
aic_table_test <- as.data.table(aic_c_table_test[2])

names(coeffs_table_test) <- c("Test", "Lambda", "q", "gamma 1", "gamma 2", "k_max")
names(aic_table_test) <- c("Test", "Poisson", "Geometric", "Zeta Gamma 2", "Zeta", "RT Zeta")


kable(coeffs_table_test)
```

And the delta AIC one:

```{r echo=FALSE}
kable(aic_table_test)
```

## Zeta test

Now we are going to apply the same tests with a prior known.

For the case of the coefficients table we get the results below:
```{r echo=FALSE}

zeta_source = read.table("./list_zeta.txt", header = TRUE, as.is = c("language","file")) # For testing 

test_table <- create_sum_table(zeta_source)
aic_c_table_test <- compute_coeffs_table(test_table)


coeffs_table_test <- as.data.table(aic_c_table_test[1])
aic_table_test <- as.data.table(aic_c_table_test[2])


names(coeffs_table_test) <- c("Test", "Lambda", "q", "gamma 1", "gamma 2", "k_max")
names(aic_table_test) <- c("Test", "Poisson", "Geometric", "Zeta Gamma 2", "Zeta", "RT Zeta")

kable(coeffs_table_test)
```
And for the delta AIC:
```{r echo=FALSE}
kable(aic_table_test)
```


## Test conclusions
Having obtained these parameters, we can now proceed to obtain the -2 log Likelihood 
for each method and compute the AIC. 

Once computed, we can produce the delta table by substracting the best AIC of each Language
from the other methods' AIC. The resulting table is the following:
```{r echo=FALSE}
kable(aic_table)
```

We can see that, in the case of the out-degree  sequence, the method what works best is
the right-truncated zeta one.


# Discussion

# Methods