---
title: "CSN - Third Lab"
author: "Kymry Burwell, Laura Cebollero"
geometry: "left=3cm,right=3cm,top=2cm,bottom=2cm"
date: "November, 2018"
output: pdf_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.pos = 'h')
```

```{r setup, include=FALSE, echo=FALSE}
wd = getwd()
if(grepl("nora", wd)) {
    setwd("~/git/csn-labs/03")
} else {
    setwd("~/Google Drive/UPC/Fall 2018/CSN/Labs/git_labs/Complex-and-Social-Networks/03")
}
rm(wd)

source("baseCode.R")
```

# Introduction

On this lab we are asked to verify the possible significance of network metrics applied
to global syntactic dependency trees, generated from large samples of 10 different languages.

The possible metrics are:

- *Clustering Coefficient* $C$ defined by Watss-Strogatz ($C_{WS}$)
- *Closeness Centrality* $\mathcal{C}$

We have chosen to work with the metric of  *Closeness Centrality* $\mathcal{C}$ and check if this metric
is significantly large when comparing the real model to other two generated randomly.

## Methodology

Our null model $H_{0}$ is each syntactic tree without chaning anything.

We are going to compare this model to two others:

- A binomial graph generated randomly. In other words, a Erdos-Renyi graph. Its number of vertices and edges will always be
the same as those on the real network.
- A randomized graph generated using the switching method.

When comparing each model to the original one, we are going to work with a "traditional" confidence interval of 0.95, so we are going to:

- Reject the model if the p-value is outside of our confidence interval, thus $$p-value > 0.05$$
- Accept the model if the p-value is inside the confidence interval, thus $$p-value \le 0.05$$

This p-value is going to be computed using the Monte-Carlo method explained during theory class.


## Data preparation
First of all, since we want to work with syntactic dependency trees, we should remove loops and
multiedges (graph's properties not present in well-defined trees structures).

Once removed, we can now proceed onto computing a summary table with:

- Language.
- N: The number of vertices.
- E: Number of edges.
- <k>: The mean degree.
- $\delta$: Network density of edges.

```{r echo=FALSE}
kable(table_1)
```

## Switching method questions
*Given two edges $u \sim v$ and $s \sim t$, what are the switchings that **

- **preserve the degree sequence?** 

- **preserve the degree sequence but produce edges that are not allowed (loops, multiedges)?**
(switchings not satisfying this property are not valid; performing them is a waste of time but they have to be counted to determine when QE switches, successful or not, have been reached)